### 2020-02-05 analysis/Paper_draft_ENG.Rmd
mas de la intro

---
title: "A Methodological Approach Comparison to Estimate Mortality in Small Areas: the Pampean Region in Argentina (2009-2011)"
author: |
  Nicolás Sacco [^ns]; Iván Williams [^iw]; Bernardo L. Queiroz [^bq]

    [^ns]: Penn State, nsacco@psu.edu
    [^iw]: Universidad Nacional de Luján, ivanwilliams1985@gmail.com
    [^bq]: Cedeplar-UFMG, lanza@cedeplar.ufmg.br

fontsize: 12pt
foonotesize: 10pt 
date: "Aug, 2019"
output: pdf_document
citation_package: natbib
bibliography: [../bib/refsubmort.bib]
link-citations: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
# change your dire
load("analysis/Rmarkdown_variables.RData")
# rmarkdown::render('Paper_draft_ESP.Rmd', encoding="UTF-8")
```

# **Abstract**

## **Background**

While the demand for epidemiological estimates of small areas is growing, recent studies show a wide gap and persistent inequalities in life expectancy at birth in Latin America. The lack of geospatial data challenge the application of different methods to study health differentials. Often data in small areas are non-existent or sparce in this region. Spatial patterns are essential to understand individual demographic outcomes related with the characteristics of a place, also as a tool for the application of development plans and for the allocation of resources.

## **Objectives**

The history of estimation in small areas is scarce in Argentina, a clear example of a country with few data sources, but also a very interesting case of the epidemiological transition, often not address by the previous literature. Based on the experience of new applications according to the available auxiliary information, in this paper we apply three different methods to estimate and then compare mortality levels in small areas in the Pampean Region of Argentina, during the period 2009-2011.

## **Methods**

Life expectancy at birht was calculated based on a Bayesian approach, a relational life-table method, smoothing spatial data ‘borrowing strength’ from neighboring areas, and third a indirect demographic approach. Estimates are compared to calculate $q_0$ rates from full death registers and then related to spatial areas, using a simple regionalization analysis.

## **Results**

Results display the potential value of the Bayesian method at the small area-spatial scale. Estimated rates indicate that there is a huge variability in life expectancy at birht withing and between regions, with a spread of up to more than 6 years in the Province of Buenos Aires. Furthermore, we find suggestive evidence that the rates by age present a higher infant mortality but also a higher risk in older adults, in provinces that exibhit more heterogeneity.

---------------------------------------------------------------------------------------------------------------------------------------

## **Introduction**

Do mortality disparities subsist among small-areas? Previous research in the demography of mortality differentials argue that they would be gradually dismiss by the epidemiological transition, public health policies, and improvements in quality of life (CITA). In most countries hig-income and middle-income countries, life expectancy at birth ($q_0$) increase during the early 20th century, due the advancement of knowledge and medical technology to fight infectious diseases and improvements in living conditions associated with socioeconomic development.

Althought socioeconomic mortality differentials contributed to sustain the hypotesis of persistent inequality at death during and after the epidemiological transition (CITA, Palloni, ?), some authors expect that health outcomes and disparities would be extinguish as the demographic change unfolds (CITA). Retrospective views, however, show that mortality differentials are persitent and have grown over time in many cases by socioeconomic status (CITA). Anyhow, it is useful account if this disparities continue and/or arise regardless of auspicious initial scenarios.

The most persuasive theoretical explanation of this issues argues that the relationship between mortality and socioeconomic position has maintained over time because the social class diferential access to information and health technology (CITA). This couse-effect hypothesis lead to significant research developments and teorical contributions. However, identify the relationshiop based on this patters and its demographic outcomes presents a serious challenge in countries where with few data sources are available. Mortality research imply the intricate patters of social, demographic and enviromental interactions. For that reason, spatial demography arise as significant perspective to answer demographic question (Matthews, 2013). In that sense, spatial patterns are essential to understand individual demographic outcomes related with the characteristics of a place.

On this subject, many recent contributions develop methods to estimate small area data. *****Mas sobre aportes recientes para estimacion de mortalidad en areas menores en el mundo.***.

Argentina represents a clear example of a country with few data sources, but also a very interesting case of the mortality transition, often not address by previous research. With all this, the aim of this article is to generate mortality estimates that exemplify the relationship between status and health. For that, we apply three different methods to estimate and then compare mortality levels in small areas in the Argentinean Pampean Region, during the period 2009-2011. To test how the choosen methods perform, we discriminate this region because is most populate region in the country.

We use administrative death register data from the Department of Health Statistics and Information of the Ministry of Health of the Nation (DEIS, acronym in Spanish; the government offical unit in charge of death counts administration), and population estimates prepare by the National Institute of Statistics and Census (INDEC, acronym in Spanish; the Argentine government agency responsible for the collection and processing of statistical data, such as population censuses). Using three different approaches, we first apply Death Distribution Methods (DDM) to analyze We then combine DDM with empirical bayesian approach to produce mortality estimates for small areas (municipalities) in Brazil. The two-step procedure, allows us to correct the underegistration of death counts and minimize random fluctuations that might occur when we calculate mortality levels in small areas.

Debido a la poca experiencia de estudios previos que estimen mortalidad *general* en áreas menores en Argentina, se decidió aplicar tres técnicas: una basada en la teoría bayesiana, la segunda basada en métodos relacionales de tablas de vida pero agregando técnicas estadísticas de suavizado, y tercero, un enfoque demográfico clásico, considerado el método por default debido a su simplicidad. Antes de la estimación, se realizó un procedimiento de regionalización para aprovechar la similitud espacial entre áreas pequeñas, independientemente de su pertenencia político-administrativa.

the quality of mortality data and estimate adult mortality levels for microregions in Brazil.

In Latin America and the Caribbean, the demand for epidemiological (and mortality, specifically) estimates of the sub-national heterogeneity is growing, both as a tool for the application of different development plans and for the allocation of resources. The experience in Latin America is led by Brazil, where there is already a sustained methodological development @FreireEtAl2015. In Argentina, experiences were found on estimates of the infant mortality trend (@torcida2008), and the communes of the Autonomous City of Buenos Aires (@Grushka2013). At the same time, more studies are focused in developing demographic tools for measuring individual heterogeneity in a population (@Vaupel_Missov_2013; @vanRaalte_Sasson_Martikainen_2018). In this sense, only considering life table information, there could be multiple layers of inequality analysis just by grouping by administrative levels.

# Why Argetina? Why the Pampas?

Debido a la poca experiencia de estudios previos que estimen mortalidad *general* en áreas menores en Argentina, se decidió aplicar tres técnicas

En Argentina, se encontraron experiencias sobre estimaciones de la tendencia de mortalidad infantil (@torcida2008) y las comunas de la Ciudad Autónoma de Buenos Aires (@Grushka2013). Al mismo tiempo, otros estudios se centran en el desarrollo de herramientas demográficas para medir la heterogeneidad en una población (@Vaupel_Missov_2013; @vanRaalte_Sasson_Martikainen_2018). En este sentido, solo considerando la información de la tabla de vida, podría haber múltiples capas de análisis de desigualdad teniendo en cuenta distintos niveles espaciales.

Este trabajo como fin tiene estimar el nivel de mortalidad para áreas menores de Argentina, en este caso, de departamentos  durante el período 2009-2011. Con el criterio de comenzar con la población con mayor participación en total país se seleccionó la región pampeana.

Se propone la siguiente línea de desarrollo. Se realiza un análisis inicial de calidad de datos, se repasan tres metodologías de suavizamiento, redefiniendo a su vez áreas mayores por fuera de lo estrictamente administrativo, se construyen tablas de mortalidad y se selecciona un método con el cual dar una primera aproximación al problema. En el medio se encuentraron inconvenientes con la información, los cuales se trataron de destacar y dar soluciones^[El material complementario, código y resultados se pueden encontrar en https://github.com/nsacco/mortalidad_Argentina.].

Este trabajo como fin tiene estimar el nivel de mortalidad para áreas menores de Argentina, en este caso, de departamentos  durante el período 2009-2011. Con el criterio de comenzar con la población con mayor participación en total país se seleccionó la región pampeana.

The lack of geospatial data challenge the application of different methods to study health differentials. Often data in small areas are non-existent or sparce in this region. Spatial patterns are essential to understand individual demographic outcomes related with the characteristics of a place, also as a tool for the application of development plans and for the allocation of resources.

En América Latina y el Caribe, la demanda de estimaciones epidemiológicas (y de mortalidad, específicamente) sobre heterogeneidad a nivel subnacional está creciendo, tanto como una herramienta para la aplicación de diferentes planes de desarrollo como para la asignación de recursos. La pregunta es siempre: ¿qué esconden los promedios regionales?

El problema principal es el de tratar con fenómenos con un pequeño número de experimentos, y en muchos casos desconocida cobertura. La experiencia en América Latina está liderada por Brasil, donde ya existe un desarrollo metodológico de avance sostenido (@FreireEtAl2015).

<!-- Algo de mortalidad general de Arg y necesidad de ver lo subnacional-->

En América Latina y el Caribe, la demanda de estimaciones epidemiológicas (y de mortalidad, específicamente) sobre heterogeneidad a nivel subnacional está creciendo, tanto como una herramienta para la aplicación de diferentes planes de desarrollo como para la asignación de recursos. La pregunta es siempre: ¿qué esconden los promedios regionales?

The objective of this work can be enunciated following these progressive steps:  
	+ Estimate the level of mortality of the departments during the period 2009-2011. With the criteria of starting with the population with bigger share to explore methodological approaches, the Pampean region was selected.  
	+ Estimate the heterogeneity between and within departments.  
	+ Test the hypothesis of the negative relationship between internal heterogeneity and the level of mortality at the province level.

In this first work, the step 1 is done ^[All the supplemental material, code and results can be found in https://github.com/nsacco/mortalidad_Argentina. It´s work in progress].

El problema principal es el de tratar con fenómenos con un pequeño número de experimentos, y en muchos casos desconocida cobertura. La experiencia en América Latina está liderada por Brasil, donde ya existe un desarrollo metodológico de avance sostenido (@FreireEtAl2015).

<!-- Algo de mortalidad general de Arg y necesidad de ver lo subnacional-->

In comparison with other Latin American countries, Argentina’s early socio-economic development, high degree of urbanization, and the expansion of formal education all influenced the reduction of mortality taking place earlier than in most other countries of the region.
This took place due to improvements in living conditions associated with socioeconomic development, rather than the advancement of knowledge and medical technology to fight infectious diseases. Although it took place more quickly and began from higher levels, the drop in mortality in Argentina can, in that sense, be likened to the pattern followed by developed countries with greater distance then from most of the rest of Latin America (Lattes, 1975; Grushka, 2014, p. 169) .

En Argentina, se encontraron experiencias sobre estimaciones de la tendencia de mortalidad infantil (@torcida2008) y las comunas de la Ciudad Autónoma de Buenos Aires (@Grushka2013). Al mismo tiempo, otros estudios se centran en el desarrollo de herramientas demográficas para medir la heterogeneidad en una población (@Vaupel_Missov_2013; @vanRaalte_Sasson_Martikainen_2018). En este sentido, solo considerando la información de la tabla de vida, podría haber múltiples capas de análisis de desigualdad teniendo en cuenta distintos niveles espaciales.

Se propone la siguiente línea de desarrollo. Se realiza un análisis inicial de calidad de datos, se repasan tres metodologías de suavizamiento, redefiniendo a su vez áreas mayores por fuera de lo estrictamente administrativo, se construyen tablas de mortalidad y se selecciona un método con el cual dar una primera aproximación al problema. En el medio se encuentraron inconvenientes con la información, los cuales se trataron de destacar y dar soluciones^[El material complementario, código y resultados se pueden encontrar en https://github.com/nsacco/mortalidad_Argentina.].

##

## **Data**

Death micro data for register years 2009, 2010 and 2011 was requested to the Department of Health Statistics and Information (DEIS, acronym in Spanish) of the Ministry of Health of the Nation. The late registration in all the country was about `r Prom_tardias`%, and the usual thought of compensation is not uniform between years and small areas, so we decided to process the database and take the registers by occurrence year (\ref{tab:def_tardias}). The percentage of unknown age was `r round(DescEdad,2)`% and unknown sex `r round(DescSexo,2)`%. The unknown department information by province is ranked in \ref{tab:SinDEP}, being `r SinDEP[1,1]` the province in worst position. This `r pUnkDepCABA`% is due to deaths that occurred in adjacent or very close departments to CABA, in the Great Buenos Aires (agglomerate in Buenos Aires province that is CABA's neighbor): Tres de Febrero (`r UnkDepCABA[1]`%), Vicente Lopez (`r UnkDepCABA[2]`%), La Matanza (`r UnkDepCABA[3]`%), Avellaneda (`r UnkDepCABA[4]`%), Lanús (`r UnkDepCABA[5]`%), Morón (`r UnkDepCABA[6]`%), San Isidro (`r UnkDepCABA[7]`%) y Gral. San Martín (`r UnkDepCABA[8]`%) ^[@Camisa_2019 noted this bias a lot of years ago: 10% of births and deaths during the period 1946-1948.]. Also because of the change in the definition of the administrative units in CABA, starting in 2011 in the data base, is not possible to match the three risk years; because of both reasons was decided leaving out of this first estimation.

The unknown data within small areas  is an important issue (\ref{tab:UnkSexAge}). Looking only the Pampean provinces, Buenos Aires had the departments with highest percent of unknown age and sex, with larger values in sex, being the leader the department `r UnkSexAge[1,5]` with `r UnkSexAge[1,6]`%. For that reason, in this first study we won't disagree the estimation by sex. The unknown categories (in variables age, residence province and residence department) were distributed proportionally.

For the exposure population for each department was used the estimated population by INDEC at middle year in 2010, and applied the census structure (@INDEC2015). Then, instead of simple averaging the three death years, was taken in account the year-person lived in the three year time window between 2009 and 2011 and the estimation date (@Gonzaga_Schmertmann_2016). This procedure allowed, first to smooth a bit the age-heaping, that could be a greater bias for rate comparison when the deaths count doesn't follow this pattern by age; and second taking in advantage of the omission corrections made by INDEC ^[It's not clear the methodology applied for the adjusted population at middle year on departments, and if takes in account a "residence" correction (@INDEC2015)]. For that was assume a uniform distribution of birth date within a year and closed population. A unique survival function was used for all the districts, applying the same standard life tables than the authors (a representative mean of the Human Mortality Database in posterior years to 1969) but weighted by a male ratio of 1.04 at birth to get both sex. A similar smooth, but with less demographic interpretation, can be achieved with local regression on 3 times the population count (@James2014) (see \ref{fig:AdjExp}).

```{r AdjExp, echo = F, fig.align='center', out.width="70%", fig.cap="Exposure Adjustment 2008-2010. Four cases. Source: own based in Census"}
include_graphics('plots/AdjExp.pdf')
```

## **Some consistency cheks**

There are increasing attempts to address the death coverage problem, depending on the auxiliary information that is counted (@Preston1980; @Bennett_Horiuchi_1984; @Schmertmann2018). The utilization of demographic indirect methods of coverage evaluation are difficult to maintain in small population very influenced by internal migration and low exposures.

With the aim to visualize possible problems of quality data in departments, two consistency checks were performed.

First, we calculated the $P/F$ method for indirect estimation of infant mortality and mapped against $q_0$ from the raw data obtained in the last section (@Moultrie). This is not an accurate method for small populations, but can give an idea about problems in biggest areas (numerator or denominator rates). We show the unweighted and weighted points to give more importance to consistency in populated areas, who was responsible for the smoothness bias in the methodological procedures described in next section. Using the mean age at birth for each province and the Latin America family of UN tables for the estimation, looking infant mortality points there is no obvious bias in the biggest areas, and no reason for exclude them (\ref{fig:PF}).

```{r PF, echo = F, fig.align='center', out.width="70%", fig.cap="Indirect estimates and raw infant mortality. Departments in selected provinces. Source: own based in Census and DEIS"}
include_graphics('plots/ChekPF.pdf')
```

Second, was mapped each department against the census poverty indicator Unsatisfied Basic Needs (NBI in Spanish, that will be introduced after), looking for some expected relationship (@Kaztman1995, @Preston_1975; @Grushka2013). In \ref{fig:NBI} are showed the NBI3 and NBI4, which measures the percentaje of school absence of children and substense incapacity in housing units (@Kaztman1995).
In color is highlighted the biggest department in the Pampean Region, called La Matanza (having `r LAMArep`% of the province Buenos Aires). La Matanza department has one of the lowest standardized mortality rate, but similar poverty index than others. Looking its performance in the poverty indexes 1 (poor housing), 2 (toilet available) and 3 (average persons by house), and having the lowest standardized mortality rate (using all region structure), it was decided to leave it out this study, where the biggest areas has the importance to smooth the small ones and this could bias the results. 
This correction is already advised in the INDEC warning of department population counts in 2010 Census, where Buenos Aires is one of the provinces with coverage issues ^[(https://www.indec.gob.ar/nivel4_default.asp?id_tema_1=2&id_tema_2=24&id_tema_3=119, visited on 25/2/19].

```{r NBI, echo = F, fig.align='center', out.width="70%", fig.cap="Standarized mortality rate and NBI. Departments in selected provinces. Source: own based in Census and DEIS"}
include_graphics('plots/ChekNBI.pdf')
```

An additional check was calculate the $e_0$ for each province using the inputs described before and the INDEC estimation (@INDEC2013). The results are a relative difference (%) of `r paste0(Dif_e0_INDEC[,4], collapse = ", ")` for provinces `r paste0(Dif_e0_INDEC[,1], collapse = ", ")`, considered an acceptable approximation given the one year of distance in temporal reference (\ref{tab:Dif_e0_INDEC}).

## **Methodology**

Due to the fact that is the first estimation in departments in Argentina (at least that is known by the authors), was decided to apply three techniques: one based in Bayesian theory, the second one based in relational life-table methods but adding statistical smoothing, and third a classical demographic approach that is considered the default method. Before the estimation, a regionalization procedure was done to take advantage of spatial similitude between small areas.

### **Regionalization**

The definition of a region or cluster must exploit the internal similarity between small areas to be able to suppose that their mortality are realizations of a greater stochastic process. The similarity in mortality patterns use to be approached  because of belonging to the same province, where the "distance" between jurisdictions is not measured by geographical distance or socio-economic attributes (@Longford2005).

To consider this we take the approach given by @AssunCao2006, which tried to define larger areas internally homogeneous and with contiguous regions in space. First is done a connectivity graph between centroids and then is calculated the cost between them (euclidean distance in our case). Then, an iteration procedure estimates the minimum spanning tree, which is the connected tree with minimum cost, measured as the sum of the dissimilarities over all the edges. Finally, a partition procedure is made cutting the edge that minimize the variance within the resulting two clusters. Because testing all the possible combinations at each partition is a computational problem, the authors proposed an heuristic approach. An overclusterization would increase homogeneity but also increase variance in smallest units because of not enough exposures. That´s the reason because of a population threshold must be settled.

As inputs for the regionalization procedure we use the shape files available online ^[https://redatam.indec.gob.ar/ , visited on 25/2/19.] and the census index Unsatisfied Basic Needs (NBI in Spanish), both by department, available in INDEC webpage ^[Same date.]. We scaled the index to standard deviation units and applied the methodology of commented before, implemented in the *spdep* package, with the function *skater* (@Bivand2019).

With this segmentation  was obtained an increase of `r AumHeterogEntreGr`% in the variance between groups and a decrease of `r DismHeterogIntraGr`% in the average variance within each group. The new cluster is more distinct between them, and with less internal relative variance (see \ref{fig:cluster}). We'll use this grouping for the calculations of two of the next three methods applied.

```{r cluster, echo = F, fig.align='center', out.width="70%", fig.cap="Regionalization of departments. Source: own based in Census and DEIS"}
include_graphics('plots/cluster.pdf')
```

### **Estimation Methods**

Although Argentina is usually classified as a country with good death statistics (@JaspersOrellana1994), from what is known there is an unequal percentage of child deaths not registered by province years ago (@DEIS2016). In this sense, since the source of data is a registry, thinking about the estimation of mortality rates by age, one might think that there is no variance, and the bias (both components of the mean squared error of an estimator) would be given by the pattern of cases omitted in each jurisdiction. As was mentioned before, the second component of the error will not be addressed in this work because of no information about its distribution by small areas (and is mentioned as a the main limitation in next sections).  Regarding the first, there are phenomena with a small number of "experiments" (few exposed in our case), which have a greater variance in their estimates, so it requires special treatment in order to reflect the risk of underlying mortality (@Brillinger1986). For achieving that we will use and compare three different methods, which have in common that take advantages of the information in the grater area.

The Empirical Bayesian method improves the statistical efficiency of the estimators of mortality rates by age, decreasing the variance in the cases of small jurisdictions (@Efron1972; @Marshall1991; @Longford2005; @Assuncao2005). The idea is that assuming that the different observations of each area proceeds from a common prior, each estimation can be improved using information from the other ones. The prior distribution corresponds to the joint distribution of the vector of mortality rates by age of the largest area. Then, through the observed behavior in each minor area, the Bayesian adjustment of the posterior mortality distribution occurs. The characteristic of "empiric" is that the distributions of the parameters of the greater area are estimated, in this case by the method of moments, also from the data.

Let's see first the univariable case. Considering a five-year age group either in an area $i$, the distribution of deaths $d$ is assumed to be a Poisson process, with expected mean $E(d_{x, 4} ^ {i}|{m_{x,4}^{i} }) = N_{x, 4}^{i} m_{x, 4}^{i}$, being $N$ the exposures and $m$ the mortality rate.

First consider $\hat{m}_{x,4}^{i} = D_{x,4}^{i}/N_{x,4}^{i}$ as the maximum likelihood estimator of the rate $m_{x,4}^{i}$ in the area $i$, which are $iid$ generated from $m_{x,4}$. The conditioned expectancy of $\hat{m}_{x,4}^{i}$ is $E_{m}(E({\hat{m}}_{x,4}^{i})/m_{x,4}^{i})=E_{m}({m_{x,4}^{i}})=m_{x,4}$ (big area rate), and conditioned variance $V({\hat{m}}_{x,4}^{i}/m_{x,4}^{i}) = \frac{m_{x,4}^{i}}{N_{x,4}^{i}}$.

The total variance of the estimator can be expressed as the sum of the variance of the $i's$ means and the expectancy of the $i's$ variances: $V_{m}(E(\hat{m}_{x,4}^{i}/m_{x,4}^{i}))  +  E_{m}(V({\hat{m}}_{x,4}^{i}/m_{x,4}^{i}))= V_{m}(\hat{m}_{x,4}^{i})  +  E_{m}(\frac{{\hat{m}}}{N_{x,4}^{i}})  =  V_{m}(m_{x,4}^{i})  +   \frac{m_{x,4}}{N_{x,4}^{i}}$. That is related to the hierarchical relation between the hyper parameter ($m_{x,4}$), the parameters ($m_{x,4}^{i}$) and its estimators $\hat{m}_{x,4}^{i}$.

The Bayesian linear estimator $\mathring{m}_{x, 4}^{i}$ which minimizes the mean squared error of ${m}_{x,4}^{i}$ (and indicators that are linear functions of this) is (Robbins, 1983):

$\mathring{m}_{x,4}^{i}=\hat{m}_{x,4}^{i}+S_{x,4}^{i}(\bar{m}_{x,4}^{i}-\hat{m}_{x,4}^{i})$

Again, it is empirical because $m_{x,4}$ is estimated by method of moments with $\bar{m}_{x,4}$, the weighted mean of small areas. The "shrinkage" factor $S_{x,4}^{i}$ is the ratio between the expectancy of the estimated variance in the small area $i$ and the unconditioned variance of the estimator, which is:

$S_{x,4}^{i}=\frac{V_{m}(m_{x,4}^{i})}{V_{m}(m_{x,4}^{i})+\frac{m_{x,4}}{N_{x,4}^{i}}}$

Or seen in another way, this formula represents the ratio between the variance of the smaller area with respect to the sum of the total variance (of the smaller and larger area), in tune with an analysis of the classic variance between groups (ANOVA). Following this reasoning, in a context of extreme homogeneity, a very small minor area could be characterized from the estimation of the largest area ($S_{x,4}^{i}\cong 1$). On the other hand, areas of high population weight will take values close to those observed ($S_{x,4}^{i}\cong 0$). In the middle of these extremes, the function linearly combines the estimate of the big area with respect to the smallest area included.

@Longford1999 extended this idea to vectors of random variables from smaller areas ("multivariate shrinkage"), estimating $S_{x,4}^{i}$ taking advantage of the correlation between sub-populations across areas. In our case, if the mortality rate of the age group between $x$ and $x + 4$ of area $i$ is greater than area $j$, a high correlation would imply that in a contiguous ages the same would happens with more chance. If covariance were null, this approach would be equivalent to univariate case described before. The calculations done in this work were done for ages 0, 1-4, and quinquennial ages until the open aged group 80. The development was done following the approach showed in @Assuncao2005 (pages 543 and 544), which estimated the parameters by the method of moments for fertility rates in Brazil.

The other method that will be applied is based in a relational mortality model called TOPALS (Tool for Projecting Age-Specific rates using Linear Splines) (@deBeer2011), which uses a linear spline to describe the ratios between the age-specific death probabilities of a given population and a standard age schedule. One advantage against the classical Brass logit approach, is that TOPALS is less sensitive by the choose of the standard. @Gonzaga_Schmertmann_2016 included the spline fit into a Poisson regression of the rates, allowing confidence intervals for the results that take in account variance for low exposure reason.

The vector of mortality rates in the small area $m^{i}(\alpha)=m^{*}* \exp^{\alpha B_{x}}$ is a function of the spline "nodes" $\alpha$, which are the ages at which will be valuated the offset from the standard,  being $m^*$ the standard mortality rate vector, and $B_{x}$ is the B-spline matrix that multiplied by $\alpha$ gives the lineal offset between the log-rates.

The idea is suppose $D_{x}\sim Poi(m_{x}N_{x})$ in each small area, construct the likelihood function using the observed deaths and exposures $\log (L(m_{x}N_{x}|D_{x}))=\sum_{\forall x}{\lbrack -m_{x}N_{x}+D_{x}\ln (m_{x})+D_{x}\ln (N_{x})-\ln (D_{x}!)\rbrack}$, but expressing that in function of the parameter $\alpha$, and adding a penalization for distance from the standard and smoothing between adjacent ages for smoothness. Using 6 nodes at ages 0, 1, 10, 20, 40, 70 and 100 like the authors, the final log-likelihood to maximize is: $Q(\alpha )=\sum_{\forall x}{\lbrack -m(\alpha )_{x}N_{x}+D_{x}\ln(m(\alpha )_{x})\rbrack }-\sum_{k=0}^{5}{(\alpha _{k}-\alpha _{k+1})^{2}}$.

Finally, as the basic method, will be applied the Indirect Standardization method, maybe one of the first approaches to this problem (@Arriaga2011). It is based in change only the level of the major area to replicate the deaths of the minor area that is being estimated. Is the case where no consideration about specific mortality shape of the minor area is made.

The Empirical Bayesian method is particularly appropriate in cases with "small local samples, substantial regional variations, strong relations between schedule components, and known spatial relationships" (@Assuncao2005). In the case of TOPALS regression, the application is more a smoothness technique than a spatial-variability model, so less assumptions about relationships between areas are needed.

### **Results**

The estimates were made for five-aged groups in all departments in Pampean Region (except La Matanza and those in CABA). The TOPALS methods was thought and applied before to simple ages, but in this case because of no correction for omission in small areas, it was decided not to increase uncertainty so was applied also in quinquennial ages with knots in ages 0, 5-9, 20-24, 40-44 and 60-64. Four examples are shown in next graph, where the differences between methods starts to appear when the exposures are smaller (see \ref{fig:Ajuste}).

```{r Ajuste, echo=F, fig.align='center', out.width="70%", fig.cap="Mortality Estimates of Departments. Different methods. Source: own based in Census and DEIS"}
include_graphics('plots/Ajuste.pdf')
```

The $e_0$ as the summary measure of mortality is considered to compare, considering also that the effect of problems in adult ages doesn´t affect hardly the estimates on $e_0$ in populations where infant, young and adult mortality still with an important weight ^[In mathematical terms, $\frac{de_0}{d\mu_{80+}}=f(T_{80+})$ are close to zero depending the non-rectangularity of $l_x$ and the correction or change in the rate; in other therms a change in these rates, for example because of bias correction, is weighted by $l_80$: $\frac{de_0}{d\epsilon}=\int_{o}^{\inf}{\frac{d\mu_x}{d\epsilon}e_x l_x dx}$ (@Wrycza2012).]

The correlation between the methods is clear: there is a strong similitude between TOPALS and Indirect estimation (`r #CorrMeth[2,1]`), but not so strong in empirical Bayes against Indirect estimation (`r #CorrMeth[3,1]`), and TOPALS (`r #CorrMeth[2,3]`) (see \ref{fig:comparativeMeth}).

```{r comparativeMeth, echo = F, fig.align='center', out.width="70%", fig.cap="Estimates of life expectancy at birth with three methods. Source: own based in Census and DEIS"}
include_graphics('plots/CompMethods.pdf')
```

The main differences are because the Bayesian method tends to take always some information about the age pattern, and assumes a correlation between them. The departments `r #cat(nombre_problems,sep = ', ')` with biggest differences are all of small populations and few non-zero cells (see \ref{fig:Feos}).

```{r Feos, echo = F, fig.align='center', out.width="70%", fig.cap="Mortality Estimates of Departments with largest differences between methods. Source: own based in Census and DEIS"}
include_graphics('plots/AjusteFeos.pdf')
```

In the next we'll take the Bayes empirical results because of the conservative scenario in ranking departments ^[A simulation study will be done here to achieve stronger conclusions]. To take randomness in account, bootstrap re-sampling of the death counts following a Poisson distribution was done. This gave us confidence intervals of the life table functions and specifically of life expectancy at birth (see \ref{fig:jerarq} and \ref{tab:table_e0}).

**An analysis: Buenos Aires**

Buenos Aires is the most populous province in Argentina, including `r n_dptos[1]` administrative areas. In the previous graph was shown his great heterogeneity, with an estimated range in life expectancy of approximately 6 years. The province can be divided between the Great Buenos Aires area made of 24 departments, an urban area that rounds CABA, and the rest of the surface. To inspect the meaningful of the results, we can take four jurisdictions with not small exposures and placed along the distribution: San Isidro, Chivilcoy, General Pueyrredón y Moreno (see \ref{fig:dptosBsAs}).

```{r dptosBsAs, echo = F, fig.align='center', out.width=c("70%"), fig.cap="Mortality Estimates of four Departments in Buenos Aires. Source: own based in Census and DEIS"}
include_graphics('plots/dptosBsAs.pdf')
```

Moreno had more infant mortality but also a greater adult and oldest risk. A priori, there is no reason to believe in a differential under count of exposures in all the age range 40+, so probably this could be a mortality pattern to consider for policymakers. In the case of San Isidro, which had the higher life expectancy at birth of this group, had the lowest hump in the typical age range of external causes. In other case, Gral. Pueyrredón had the worst position in the age range 5 to 25, ages that pushed up mortality for preventable causes.

##**Future work**

+ To decide which one of these method could be use to test the heterogeneity hypothesis, one option is simulate known mortality profiles in different scales and patterns of omissions.

+ Some additional consistencies have to be made between the posterior distribution in the Bayes Empirical method and the official estimates at the province level, to guarantee similar levels.

+ We analyzed two layers of heterogeneity: between provinces (considering variance of departments $e_0$) and from a life table perspective, looking for inequality in the death distribution between departments in each province. Still thinking in the province level is important because the administrative areas are determinant in the health policy strategies for each government.

##**Limitations**

The level of coverage of minor areas is unknown. Some visual checks were made on the consistency of the mortality level of the department with respect to socio-economic indicators, in order to detect possible anomalies, but just focusing in the grater departments because of the statistical properties of the methods.

##**Conclusions**

The demographic research, like other social sciences, is limited by the available and quality source of information. In general, when the information is not of good quality or just very expensive to collect, the demography field built a big history in applying indirect methods to start to get closer to real values of the demographic phenomena. This work goes in this direction. With no background (at least that We know) some things can be said about the mortality in small areas in Argentina. We decided to start with Pampean Region because of the population share of the country. We applied three methods to estimate mortality structure and level, and We made previous  consistency checks to discard big problems. The main differences are because the Bayesian method tends to take always some information about the age pattern, assuming correlation between them. One important conclusion was made for the Buenos Aires case, characterizing 4 departments and quantifying different mortality profiles. This can be a start point for give priorities in health policies and be focused by future research.

##**References**

## Tables

```{r SinDEP, echo = F, results='asis', error=F, warning=F}
library(kableExtra)
kable(SinDEP, digits = 2, caption = 'Provinces with unknown residence department',
      col.names = c("Province", 'Unknown %')) %>% #'$Ls(IC95)$',
      add_footnote(c("Source: Own calculations based in DEIS registers"))
```

```{r UnkSexAge, echo = F, results='asis', error=F, warning=F}
kable(UnkSexAge, digits = 2, caption = 'Dptos sin dato', row.names = F) %>%
      add_footnote(c("Source: own calculations based in DEIS registers"))
```

```{r def_tardias, echo = F, results='asis', error=F, warning=F}
kable(def_tardias, caption='Registered and Ocurred. Argentina') %>%
      add_footnote(c("The year register is in rows, and ocurred in columns. Source: own calculations based in DEIS registers"))  %>%
      add_header_above(c(" " = 1, "Ocurred year" = 6))
```

```{r Dif_e0_INDEC, echo = F, results='asis', error=F, warning=F}
library(kableExtra)
kable(Dif_e0_INDEC, digits = 2, caption='Differences between life expectancy calculated with non adjusted data and official estimates',
      col.names = c('Province', "Own", "INDEC", 'Realtive Diference')) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
      full_width = F, position = "center")%>%
      add_footnote(c("Source: Own calculations and INDEC (2013)"))
```

```{r table_e0, echo = F, results='asis', error=F, warning=F}
library(kableExtra)
kable(table_e0, digits = 2, caption='Life expectancy at birth estimates. 2008-2010',
      col.names = c('Province', "Department", "Mean", 'p2.5', 'p97.5')) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
      full_width = F, position = "center")%>%
      add_footnote(c("Source: Own calculations"))
```